---
title: "grade_analysis"
author: "Disem Sula"
format: html
editor: visual
---

```{r}
#| echo: false
#| message: false
here::i_am("grade_analysis_LAB.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(stringr)
```

## Grade analisys LAB

### Question 1

```{r}
#| message: false
grades <- vroom(here("grades.csv"))
```

### Question 2

```{r}
#| echo: false
#| message: false
grades |>
  summarise(min(Exam, na.rm = TRUE), max(Exam, na.rm = TRUE), median(Exam, na.rm=TRUE), mean(Exam,na.rm = TRUE),n()) |>
  knitr::kable()
```

### Question 3

```{r}
#| echo: false
#| message: false
missingstud<-sum(is.na(grades$Exam))
```

There is a total of `r missingstud` students who passed the test

### Question 4

```{r}
#| echo: false
#| message: false
ggplot(grades, aes(x=Exam))+
  geom_histogram(binwidth=1.5, na.rm=TRUE) +
  labs(x="Exam grades", y= "frequency")
```

### Question 5

```{r}
#| echo: false
#| message: false
grades |>
  group_by(Group) |>
  summarise(N=n()) |>
  knitr::kable()
```

### Question 6

```{r}
#| echo: false
#| message: false
group_counts <- grades |>
  group_by(Group) |>
  summarise(N = n())

ggplot(group_counts, aes(x = Group, y = N)) +
  geom_bar(stat = "identity") +
  labs(x = "Group", y = "Number of Students") +
  ggtitle("Number of Students in Each Group") +
  theme(axis.text.x = element_text(angle = 90))
```

### Question 7

```{r}
#| echo: false
#| message: false
ggplot(grades, aes(x = Group, y = Exam)) + 
  geom_bar(stat = "identity", position = "dodge", na.rm = TRUE) +
  labs(x="Group", y="Exam Grades", title = "Distribution of Grades at the Exam by Group") +
  theme(axis.text.x = element_text(angle = 90))
```

In the following representation we use the function geom_boxplot, that way we can see the median, quartiles and possible outliers

```{r}
#| echo: false
#| message: false
ggplot(grades, aes(x = Group, y = Exam)) +
  geom_boxplot(na.rm = TRUE) +
  labs(x = "Group", y = "Exam Grades", title = "Distribution of Grades at the Exam by Group") +
  theme(axis.text.x = element_text(angle = 90))
```

### Question 8

```{r}
#| echo: false
#| message: false
group_missed <- grades |>
  group_by(Group) |>
  summarise(Missed_Exam = sum(is.na(Exam)))
group_missed |>
  knitr::kable()
```

### Question 9

```{r}
#| echo: false
#| message: false
ggplot(group_missed, aes(x=Group, y=Missed_Exam)) +
  geom_col() +
  labs(x = "Group", y = "# of Students Who Missed Exam", title = "Students Who Missed the Exam in Each Group") +
  theme(axis.text.x = element_text(angle = 90))
```

### Question 10

```{r}
#| message: false
grades_long <- grades |>
  pivot_longer(cols = starts_with(("MCQ_")),
    names_to = "name",
    values_to = "value") |>
  select(Id, Group, name, value)
```

### Question 11

```{r}
#| message: false
missing_grades <- grades_long |>
  group_by(Id) |>
  summarise(Missing_by_stud = sum(is.na(value)))
```

### Question 12

```{r}
#| echo: false
#| message: false
ggplot(missing_grades, aes(x=Missing_by_stud)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Number of Missing Grades", y = "Frequency", title = "Distribution of Missing Grades per Student")
```

### Question 13

```{r}
#| echo: false
#| message: false
group_missed_long <- grades_long |>
  filter(str_starts(Group, "grp_1")) |>
  group_by(Group) |>
  summarise(Students_Missed_Exam = sum(is.na(value)))
group_missed_long |>
  knitr::kable()
```

### Question 14

For this question I created a new data.frame as in question 10, but with online exams, then I computed the number of missing grades for the online tests for each student

```{r}
online_grades_long <- grades |>
  pivot_longer(cols = starts_with(("Online_MCQ_")),
    names_to = "name",
    values_to = "value") |>
  select(Id, Group, name, value)

missing_online_tests <- online_grades_long |>
  filter(str_starts(Group, "grp_1")) |>
  group_by(Id) |>
  summarise(Missing_Online_by_stud=sum(is.na(value)))
```

### Question 15

```{r}
#| echo: false
#| message: false
grades_with_missing_online <- grades |>
  left_join(missing_online_tests, by = "Id") |>
  select(Id, Exam, starts_with("Online_MCQ"), Missing_Online_by_stud)


ggplot(grades_with_missing_online, aes(x =Exam, y = Missing_Online_by_stud)) +
  geom_point(na.rm = TRUE) +
  labs(x = "Exam Grades", y = "# of Missing Online Tests", title = "Distribution of Grades at the Exam conditioned by Missing Online Tests")
```

### Question 16

```{r}
#| message: false
missed_table <- grades |>
  rowwise() |>
  mutate(Missed = any(is.na(c_across(starts_with("MCQ_"))))) |>
  select(Id, Missed) |>
  distinct()
```

### Question 17

```{r}
#| message: false
#| echo: false
group_missed_table <- grades |>
  rowwise() |>
  mutate(Missed = any(is.na(c_across(starts_with("MCQ_"))))) |>
  select(Id, Group, Missed) |>
  distinct() |>
  group_by(Group) |>
  summarise(P_missed =(sum(Missed) / n() * 100))

group_missed_table |>
  knitr::kable()
```

### Question 18

```{r}
#| message: false
#| echo: false
avg_grade_per_group <- grades |>
  group_by(Group) |>
  summarise(Avg_Exam_Grade = mean(Exam, na.rm = TRUE))

merged_table <- inner_join(group_missed_table, avg_grade_per_group, by = "Group")

ggplot(merged_table, aes(x = P_missed, y = Avg_Exam_Grade)) +
  geom_point() +
  labs(x = "Percentage of Missed by group", y = "Average  Grade", title = "Avg Grade by Group vs. % Missed exam") +
  theme_minimal()
```
